---
title: "Take-home Ex2"
author: "WuYanjie"
---

Install package
```{r}
pacman::p_load(tidyverse,readxl,corrplot,data.table,lubridate,ggthemes,gridExtra,
  scales,viridis,knitr,ggridges, patchwork)
```

import data
```{r}
data <- read_excel("data/data.xlsx")
```

Data preparation
```{r}
df_clean <- data %>%
  mutate(date = make_date(year, month, 1)) %>%
  arrange(date)

df_wide <- df_clean %>%
  pivot_wider(names_from = category, values_from = total) %>%
  mutate(
    trade_balance = output - input,  #贸易顺差
    trade_total   = output + input,  #贸易总额
    year          = year(date)
        )
```

EDA

1.total trade
2.import vs export trend
3.trade balance
4.top countries
5.country-level structure
6.country-level input vs output trend
7.country-level trade balance trend
8.monthly seasonal pattern
9 seasonal heatmap
10.seasonal ribbon plot

```{r}
#1 Overall Trade Total Trend
df_total <- df_clean %>%
  group_by(date) %>%
  summarise(trade_total = sum(total, na.rm = TRUE))

ggplot(df_total,
       aes(x = date, y = trade_total)) +
  geom_line(color = "steelblue", linewidth = 1) +
  labs(
    title = "Total Trade Over Time",
    x = "Date",
    y = "Trade Total"
  ) +
  theme_minimal()
```


```{r}
# 2 input output trend
df_clean %>%
  group_by(date, category) %>%
  summarise(total = sum(total)) %>%
  ggplot(aes(date, total, color = category)) +
  geom_line()
```

```{r}
#3 trade_balance
df_wide %>%
  group_by(date) %>%
  summarise(balance = sum(trade_balance)) %>%
  ggplot(aes(date, balance)) +
  geom_line()
```


```{r}
#4 top trading partners
df_clean %>%
  group_by(country) %>%
  summarise(total_trade = sum(total)) %>%
  arrange(desc(total_trade))
```

```{r}
#5 trade country-level structure
ggplot(df_clean,
       aes(country, total, fill = category)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Trade Country-level Structure",
    x = "Country",
    y = "Trade Value"
  ) +
  theme_minimal()
```

```{r}
#6 country-level input vs output trend
ggplot(df_clean, aes(x = date, y = total, color = country)) +
  geom_line(alpha = 0.7) +
  facet_wrap(~category, scales = "free_y") +
  theme_minimal() +
  labs(title = "Singapore Trade Trend (2000-2026)", y = "Total Value")
```


```{r}
#7 country-level trade balance trend

df_wide_plot <- df_wide %>%
  mutate(month = factor(month(date), levels = 1:12)) 

ggplot(df_wide_plot, aes(x = month, y = trade_balance, group = year)) +
  geom_line(aes(color = year), alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~country) +
  scale_color_viridis_c() +
  theme_tufte() +
  labs(title = "Monthly Trade Balance Seasonality",
       subtitle = "Output - Input (Red line indicates balanced trade)",
       x = "Month", y = "Balance Amount")
```

```{r}
#8 Monthly seasonal pattern

season_month <- df_clean %>%
  mutate(month = factor(month(date), levels = 1:12)) %>%
  group_by(month, category) %>%
  summarise(avg_trade = mean(total, na.rm = TRUE))

ggplot(season_month,
       aes(month, avg_trade, color = category, group = category)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Seasonal Pattern of Import and Export",
    x = "Month",
    y = "Average Trade Value"
  ) +
  theme_minimal()
```


```{r}
#9 seasonal heatmap
season_heat <- df_clean %>%
  mutate(month = factor(month(date), levels = 1:12)) %>%
  group_by(year, month, category) %>%
  summarise(total = sum(total, na.rm = TRUE))
ggplot(season_heat,
       aes(month, year, fill = total)) +
  geom_tile() +
  facet_wrap(~category) +
  scale_fill_viridis_c() +
  labs(
    title = "Seasonal Trade Heatmap",
    x = "Month",
    y = "Year"
  ) +
  theme_minimal()
```

```{r}
#10 seasonal ribbon plot

season_ribbon <- df_clean %>%
  mutate(month = factor(month(date), levels = 1:12, labels = month.abb)) %>%
  group_by(month, category) %>%
  summarise(
    avg_trade = mean(total, na.rm = TRUE),
    min_trade = min(total, na.rm = TRUE),
    max_trade = max(total, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(season_ribbon,
       aes(x = month, y = avg_trade,
           color = category, fill = category,
           group = category)) +
  geom_ribbon(aes(ymin = min_trade, ymax = max_trade),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Seasonal Trade Pattern with Variation Range",
    x = "Month",
    y = "Average Trade Value"
  ) +
  theme_minimal()

```




CDA 


1.Import 和 Export 是否显著不同
2.不同国家的贸易规模是否不同
3.贸易顺差是否随时间变化
4.国家之间贸易结构是否不同
5.Trade values vary significantly across months




```{r}
ggplot(data, aes(total)) +
  geom_histogram(bins = 30,
                 fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Distribution of Trade Values",
    x = "Trade Value",
    y = "Frequency"
  )
```



```{r}
#H0: Import and export values are the same
#H1: Import and export values differ
ggplot(df_clean,
       aes(category, total, fill = category)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.1) +
  labs(
    title = "Distribution of Import and Export Values",
    x = "Trade Category",
    y = "Trade Value"
  ) +
  theme_minimal()

```

```{r}
#H0: Trade values are equal across countries
#H1: Trade values differ across countries

ggplot(df_clean,
       aes(reorder(country, total), total)) +
  geom_boxplot(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Trade Value Distribution by Country",
    x = "Country",
    y = "Trade Value"
  ) +
  theme_minimal()
```

```{r}
#Trade balance 是否随时间变化
balance_year <- df_wide %>%
  group_by(year) %>%
  summarise(trade_balance = sum(trade_balance, na.rm = TRUE), .groups = "drop")

ggplot(balance_year, aes(x = year, y = trade_balance)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Trend of Trade Balance Over Time",
    x = "Year",
    y = "Trade Balance"
  ) +
  theme_minimal()
```

```{r}
#国家之间贸易结构是否不同

df_wide <- df_wide %>%
  mutate(export_ratio = output / trade_total)

anova_ratio <- aov(export_ratio ~ country, data = df_wide)
summary(anova_ratio)

ggplot(df_wide, aes(country, export_ratio)) +
  geom_boxplot() +
  coord_flip()
```

```{r}
#Trade values vary significantly across months

anova_month <- aov(trade_total ~ factor(month), data = df_wide)
summary(anova_month)

ggplot(df_wide, aes(factor(month), trade_total)) +
  geom_boxplot()
```





