---
title: "Hands-on——Ex07"
author: "Wu Yanjie"
---

# **Visualising and Analysing Time-oriented Data**

## **Getting Started**

### **Install R Package**

```{r}
pacman::p_load(scales,viridis,lubridate,ggthemes,
               gridExtra,readxl,knitr,data.table, 
               CGPfunctions, ggHoriPlot, tidyverse)
```

### **Importing the data**

```{r}
attacks <- read_csv("data/eventlog.csv")
```

### Examining the data structure

```{r}
kable(head(attacks))
```

There are three columns, namely timestamp, source_country and tz.

timestamp field stores date-time values in POSIXct format. source_country field stores the source of the attack. It is in ISO 3166-1 alpha-2 country code. tz field stores time zone of the source IP address.

### Data Preparation 

Step1: Deriving weekday and hour of dayfields

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

Step2: Deriving the attacks tibble data frame

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

Beside extracting the necessary data into attacks data frame, mutate() of dplyr package is used to convert wkday and hour fields into factor so they’ll be ordered when plotting. Table below shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

Building the Calendar Heatmaps

```{r}
grouped <- attacks %>%
  count(wkday, hour) %>%
  ungroup() %>%
  na.omit()

ggplot(grouped,
       aes(hour,
           wkday,
           fill = n)) +
geom_tile(color = "white",
          size = 0.1) +
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue",
                    high = "dark blue") +
labs(x = NULL,
     y = NULL,
     title = "Attacks by weekday and time of day")
theme(axis.title = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6))
```

a tibble data table called grouped is derived by aggregating the attack by wkday and hour fields. a new field called n is derived by using group_by() and count() functions. na.omit() is used to exclude missing value. geom_tile() is used to plot tiles (grids) at each x and y position. color and size arguments are used to specify the border color and line size of the tiles. theme_tufte() of ggthemes package is used to remove unnecessary chart junk. To learn which visual components of default ggplot2 have been excluded, you are encouraged to comment out this line to examine the default plot. coord_equal() is used to ensure the plot will have an aspect ratio of 1:1. scale_fill_gradient() function is used to creates a two colour gradient (low-high).

### Building Multiple Calendar Heatmaps

Step 1: Deriving attack by country object Following will be done to indentify the top4 countries with the highest number of attacks: count the number of attacks by country calculate the percent of attacks by country save the results in a tibble data frame

```{r}
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```

Step 2: Preparing the tidy data frame

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

### Plotting Multiple Calendar Heatmaps

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r}
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

### Plotting Cycle Plot

**Step 1: Data Import**

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

**Step 2: Deriving month and year fields**

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

**Step 3: Extracting the target country**

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

**Step 4: Computing year average arrivals by month**

```{R}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

**Srep 5: Plotting the cycle plot**

```{r}
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

### Plotting Slopegraph

**Step 1: Data Import**

```{r}
rice <- read_csv("data/rice.csv")
```

**Step 2: Plotting the slopegraph**

```{r}
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Dr. Kam Tin Seong")
```
