---
title: "Take-home_Ex01"
author: "Wu Yanjie"
---

# 1. Introduction

## 1.1 Background

Academic performance is shaped by a combination of personal characteristics, learning behaviors, and institutional experiences. However, these factors do not necessarily influence performance by creating distinct groups of high- and low-achieving students; instead, they may gradually reposition students along a performance continuum. Understanding whether variables produce clear separation or merely shift distributional placement is therefore essential for interpreting their practical relevance in educational settings.

This study examines survey responses from students at the University of Education, Vietnam National University, Hanoi. Rather than focusing on prediction, the analysis investigates how different factors are structurally associated with ordered GPA categories, emphasizing distributional patterns instead of mean differences.

## 1.2 Objectives

The study aims to evaluate how academic performance varies across three domains:

Demographic background (gender, economic status)

Behavioral engagement (study time, social media use)

Institutional experience (learning adaptation, lecturer support and quality)

By comparing distributional structures across ordered categories, the analysis distinguishes between variables that create systematic gradients and those that produce only minor compositional differences. The goal is not to infer causality but to identify how different factors align with students’ relative position within the performance hierarchy.

# 2. Getting Started

## 2.1 Load Packages

We load the following R packages using the pacman::p_load() function to support data import, wrangling, visualisation, and statistical graphics: tidyverse: Core collection of packages for data wrangling (dplyr), data import (readr), and data visualisation (ggplot2). scales: Provides scaling functions (e.g., percentage labels) used in ggplot2 visualisations. ggthemes: Offers additional themes to improve plot aesthetics beyond default ggplot2 themes. ggridges: Used to create ridgeline density plots for comparing GPA distributions across ordinal groups. ggdist: Provides visualisations of distributions and uncertainty (e.g., median + interval plots using stat_pointinterval()). ggstatsplot: Produces enhanced statistical visualisations such as ggbetweenstats(), integrating non-parametric tests and effect sizes directly into the plots. colorspace: Supports colour palettes and colour manipulation utilities (useful when fine-tuning plot appearance).

```{r}
pacman::p_load(tidyverse, ggdist, ggridges, ggthemes, scales, ggstatsplot, colorspace)
```

## 2.2 Import data

The data for this exercise is retrieved from the publicly available survey dataset published on Mendeley Data (https://data.mendeley.com/datasets/23ppcdbmhc/1)

```{r}
#| message: false
#| warning: false
Database_paper <- read_csv("data/Database_paper.csv")
```

The dataset contains 2,170 survey responses collected in 2023.\

## 2.3 Variable Selection

In order to investigate factors associated with students’ academic performance (GPA), a subset of relevant variables was selected from the survey dataset based on theoretical relevance and analytical purpose.

The dependent variable in this study is GPA, treated as an ordered categorical variable ranging from Poor to Excellent.

The independent variables were selected across three broad dimensions:

Demographic factors (Gender, Poor household status)

Behavioral factors (Study time, Social media time)

Institutional factors (Learning adaptation, Lecturer support)

These variables were chosen because prior educational research suggests that academic outcomes are influenced by a combination of personal background, study behaviors, and institutional context.

Details of the selected variables are extracted from the survey Codebook and summarized in Table below.

```{r}
#| echo: false
#| message: false
#| warning: false
library(knitr)

var_table <- data.frame(
  Category = c(
    "Academic outcome",
    "Demographic",
    "Demographic",
    "Behavioral",
    "Behavioral",
    "Institutional",
    "Institutional"
  ),
  
  Variable_Name = c(
    "GPA",
    "Gender",
    "Poor_Stu",
    "Time_Studying",
    "Time_SocicalMedia",
    "Adapt_Learning_Uni",
    "SupportOf_Lec"
  ),
  
  Description = c(
    "Cumulative Grade Point Average",
    "Student gender",
    "Whether student belongs to poor household",
    "Average daily study time",
    "Average daily social media usage time",
    "Level of adaptation to university learning environment",
    "Level of lecturer support"
  ),
  
  
  Encoding = c(
    "1=Poor; 2=Average; 3=Fair; 4=Good; 5=Excellent",
    "1 = Male; 2 = Female",
    "1 = Yes; 2 = No",
    "1=<2h; 2=1~2h; 3=2~3h; 4=3~4h; 5=>4h",
    "1 = <1h ... 5 = >4h",
    "1=Not at all; 2=Little; 3=Moderate; 4=Quite; 5=Very",
    "1=Not at all; 2=Little; 3=Moderate; 4=Quite; 5=Very"
  )
)

kable(var_table, caption = "Table 1. Selected Variables and Their Coding Scheme")

```

We standardised the dataset by converting key variables into ordered factors with clear labels (1–5) for ordinal survey items, and recoding nominal variables (e.g., Gender, Poor_Stu) into descriptive categories to support consistent plotting and comparison.

```{r}
df <- Database_paper %>%
  mutate(
    GPA_num = as.numeric(as.character(GPA)),
    GPA_f = factor(
      GPA, 
      levels = 1:5, 
      labels = c("Poor", "Average", "Fair", "Good", "Excellent"),
      ordered = TRUE),
    Time_Studying      = factor(
      Time_Studying,      
      levels = 1:5, 
      labels = c("under 1h", "from 1 to lower 2h", "from 2 to lower 3h", "from 3 to lower 4h", "over 4h"),
      ordered = TRUE),
    Time_SocicalMedia  = factor(
      Time_SocicalMedia,  
      levels = 1:5, 
      labels = c("under 1h", "from 1 to lower 2h", "from 2 to lower 3h", "from 3 to lower 4h", "over 4h"),
      ordered = TRUE),
    Time_Friends       = factor(Time_Friends,       levels = 1:5, ordered = TRUE),
    Adapt_Learning_Uni = factor(
      Adapt_Learning_Uni, 
      levels = 1:5, 
      labels = c("Not at all", "Little", "Moderate", "Quite", "Very"),
      ordered = TRUE),
    SupportOf_Lec      = factor(
      SupportOf_Lec,      
      levels = 1:5, 
      labels = c("Not at all", "Little", "Moderate", "Quite", "Very"),
      ordered = TRUE),
    Quality_Lecturer   = factor(
      Quality_Lecturer,   
      levels = 1:5, 
      labels = c("Not at all", "Little", "Moderate", "Quite", "Very"),
      ordered = TRUE),
    Competitive_Class  = factor(Competitive_Class,  levels = 1:5, ordered = TRUE),
    Gender = factor(Gender, labels = c("Male", "Female")),
    Poor_Stu = factor(Poor_Stu,
                      labels = c("Poor household", "Non-poor household")),
    Year   = factor(Year, levels = 1:5, ordered = TRUE)
  )
```

All ordinal survey variables were converted into ordered factors. GPA was treated as an ordered categorical variable. Numeric coding was used solely to facilitate rank-based visualization and non-parametric comparison.

Although the primary objective of this study is exploratory, non-parametric group comparisons (Kruskal–Wallis tests) are conducted where appropriate to assess whether observed differences across ordinal categories are statistically meaningful. Given that GPA and most explanatory variables are measured on ordinal scales, rank-based methods are preferred over parametric alternatives. Statistical results are interpreted alongside effect sizes to emphasize practical rather than merely statistical significance.

# 3. Overall GPA Distribution

To understand the general academic performance of students, we first examine the overall distribution of GPA categories. GPA is treated as an ordered categorical variable ranging from “Poor” to “Excellent.” Since the objective is exploratory rather than inferential, we focus on visualizing how GPA values are distributed across the student population. We use a bar chart to display the proportion of students in each GPA category. This allows clear comparison of relative frequencies and provides an intuitive overview of central tendency and dispersion. Percentages are used instead of counts to improve interpretability.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(x = GPA_f)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Overall Distribution of GPA",
    x = "GPA",
    y = "Percentage of Students")
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(x = GPA_f)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Overall Distribution of GPA",
    x = "GPA Category",
    y = "Percentage of Students"
  )

```
:::

The overall distribution of GPA categories reveals that the majority of students are concentrated in the “Fair” and “Good” performance levels, with noticeably fewer students in the “Poor” and “Excellent” categories. This indicates that academic outcomes within the sample are moderately centralized rather than highly polarized. The relatively small proportion of students at the extremes suggests limited academic disparity at both ends of the performance spectrum. The clustering around mid-level categories may reflect common institutional grading structures or similar academic engagement levels among students. Establishing this baseline distribution is important because it provides a reference framework for interpreting subgroup differences in subsequent analyses. Without understanding the central tendency and spread of GPA levels, it would be difficult to meaningfully evaluate whether behavioral or institutional factors correspond with substantial performance shifts.

# 4. Demographic Factors

## 4.1 Gender and GPA

To examine whether demographic characteristics influence academic outcomes, we compare GPA distribution across gender groups. Gender is treated as a categorical variable with two levels: Male and Female. GPA remains an ordered categorical variable.

We use a grouped bar chart showing the percentage distribution of GPA within each gender. Displaying proportions rather than raw counts allows fair comparison across groups regardless of group size. The side-by-side bars facilitate visual comparison of performance patterns.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(x = GPA_f, fill = Gender)) +
  geom_bar(aes(y = after_stat(prop), group = Gender),
           position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "GPA Distribution by Gender",
    x = "GPA",
    y = "Percentage"
  )
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(x = GPA_f, fill = Gender)) +
  geom_bar(aes(y = after_stat(prop), group = Gender),
           position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "GPA Distribution by Gender",
    x = "GPA",
    y = "Percentage"
  )
```
:::

::: callout-tip
## Observation

The GPA distributions for male and female students show substantial overlap across all performance categories. Both groups are mainly concentrated in the “Fair” and “Good” levels, and no consistent directional shift across the ordered GPA scale is observed. Female students have a slightly higher proportion in the “Excellent” category, while males appear marginally more represented in lower categories; however, these differences occur locally rather than progressively across adjacent levels. The absence of a gradient pattern suggests distributional similarity rather than a systematic association between gender and academic performance. Overall, gender does not appear to reorganize the performance hierarchy but only introduces minor fluctuations within categories, indicating that demographic grouping plays a limited role compared with behavioral or institutional factors examined later.
:::

## 4.2 Family Economic Status and Academic Performance

Family economic background was proxied using the binary survey indicator Poor_Stu (Poor household vs Non-poor household). GPA distributions were compared between the two groups using grouped percentage bar charts. Proportions were computed within each economic category to ensure meaningful structural comparison. This visually driven approach enables assessment of distributional shifts without assuming numeric distance between GPA categories.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(x = GPA_f, fill = factor(Poor_Stu))) +
  geom_bar(aes(y = after_stat(prop), group = factor(Poor_Stu)), position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "GPA Distribution by Family Economic Status", x = "GPA", y = "Percentage") +
  theme_minimal()
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(x = GPA_f, fill = factor(Poor_Stu))) +
  geom_bar(aes(y = after_stat(prop), group = factor(Poor_Stu)), position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "GPA Distribution by Family Economic Status", x = "GPA", y = "Percentage") +
  theme_minimal()
```
:::

::: callout-tip
## Observation

Students from non-poor households show a slightly higher proportion in the “Good” and “Excellent” GPA categories, while students from poor households are more concentrated in the “Fair” level. However, these differences do not develop progressively across adjacent performance categories and substantial overlap remains across the full distribution. The pattern therefore reflects a modest shift in concentration rather than a systematic redistribution of performance positions. Unlike behavioral variables, economic background does not produce a clear ordered gradient along the GPA scale but instead introduces a small imbalance around the central categories. This suggests that family economic status is associated with variation in performance composition, yet the association is weak in structure and does not substantially reorganize the academic performance hierarchy.
:::

# 5. Behavioral Factors

## 5.1 Study Time and GPA Composition

To examine how daily study duration relates to academic performance, we compare the proportional distribution of GPA categories across study-time groups. Study time is treated as an ordered categorical variable representing increasing hours of daily study, while GPA is recoded into ordered categories ranging from “Poor” to “Excellent.”

A 100% stacked bar chart is used to visualize the composition of GPA within each study-time group. This approach allows direct comparison of the relative share of performance categories across groups, independent of sample size differences. By standardizing each bar to 100%, we focus on structural differences in performance distribution rather than absolute counts.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(y = Time_Studying, fill = GPA_f)) +
  geom_bar(position = "fill", width = 0.65) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title = "GPA Composition by Study Time",
    x = "Percentage",
    y = NULL,
    fill = "GPA Category"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(y = Time_Studying, fill = GPA_f)) +
  geom_bar(position = "fill", width = 0.65) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title = "GPA Composition by Study Time",
    x = "Percentage",
    y = NULL,
    fill = "GPA Category"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
:::

::: callout-tip
## Observation

A clear directional pattern emerges across study-time categories. As daily study duration increases, the proportion of students in the “Good” and “Excellent” categories rises, while lower performance categories gradually decline. Importantly, this shift occurs progressively across adjacent levels rather than appearing only at extreme groups, indicating a systematic redistribution of performance rather than isolated differences. Although overlap between groups remains, the ordered gradient suggests a structured association between study engagement and academic outcomes. Compared with demographic factors, study time shows a more coherent pattern across the ordinal scale, implying that variation in performance is aligned with behavioral intensity rather than categorical group membership. This relationship reflects a consistent performance shift rather than a sharp separation between groups..
:::

## 5.2 Social Media Usage and GPA

Daily social media usage was examined as a behavioral variable potentially associated with academic performance. GPA category distributions were compared across usage-level groups using grouped percentage bar charts. Proportions were calculated within each social media category to allow structural comparison without assuming interval distance between GPA levels.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(x = Time_SocicalMedia, y = GPA_num, fill = Time_SocicalMedia)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  scale_fill_brewer(palette = "Blues") +
  geom_jitter(width = 0.3, alpha = 0.1) +
  labs(title = "GPA by Social Media Time",
       x = "Daily Social Media Time",
       y = "GPA Level") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 16)
  )
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(x = Time_SocicalMedia, y = GPA_num, fill = Time_SocicalMedia)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  scale_fill_brewer(palette = "Blues") +
  geom_jitter(width = 0.3, alpha = 0.1) +
  labs(title = "GPA by Social Media Time",
       x = "Daily Social Media Time",
       y = "GPA Level") +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 16)
  )
```
:::

::: callout-tip
## Observation

The GPA distributions across social media usage levels show substantial overlap and lack a consistent directional pattern. Minor proportional fluctuations appear between categories; however, these differences do not progress systematically across adjacent usage levels. Both low- and high-usage groups remain concentrated in the “Fair” and “Good” performance categories, and higher usage does not correspond to a steady movement toward either higher or lower GPA levels. The absence of an ordered gradient suggests that social media time does not reorganize the performance distribution but only introduces small local variations. Compared with study time, the relationship appears irregular rather than structured, indicating that daily social media usage alone provides limited differentiation in academic performance within this dataset.
:::

# 6. Institutional Factors

## 6.1 Learning Adaptation and GPA

Students’ adaptation to the university learning environment was examined as an institutional factor potentially associated with academic performance. GPA was treated as an ordered categorical variable, and proportions were calculated within each adaptation level to allow fair structural comparison. A visually driven grouped percentage bar chart was used to evaluate distributional shifts across adaptation categories.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggplot(df, aes(x = Adapt_Learning_Uni, y = GPA_num)) +
  stat_pointinterval(
  point_interval = median_qi,
  .width = 0.95
  ) +
  labs(title = "GPA by Learning Adaptation Level")
```

### Code

```{r}
#| echo: true
#| eval: false
ggplot(df, aes(x = Adapt_Learning_Uni, y = GPA_num)) +
  stat_pointinterval(
  point_interval = median_qi,
  .width = 0.95
  ) +
  labs(title = "GPA by Learning Adaptation Level")
```
:::

::: callout-tip
## Observation

A consistent upward gradient is observed across learning adaptation levels. Students reporting higher adaptation show progressively greater representation in the “Good” and “Excellent” categories, accompanied by a reduction in lower performance levels. Unlike isolated category differences, this shift occurs across multiple adjacent adaptation levels, indicating a structured redistribution of academic outcomes along the ordinal scale. Although group overlap remains, the monotonic pattern suggests alignment between adjustment to the learning environment and performance positioning. Compared with demographic variables, adaptation demonstrates a clearer ordering effect, where higher adjustment corresponds to systematically higher performance placement rather than sporadic differences. This pattern reflects a coherent association across the performance hierarchy rather than discrete group separation.
:::

## 6.2 Lecturer Quality and GPA

Perceived lecturer quality was analyzed as an institutional variable reflecting instructional environment. GPA distributions were compared across quality perception levels using grouped percentage bar charts, with proportions calculated within each category to ensure comparability. This visually driven approach enables evaluation of structural differences without assuming interval-scale properties.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggbetweenstats(
  data = df,
  x = Quality_Lecturer,
  y = GPA_num,
  type = "np",
  messages = FALSE,
  title = "Lecturer Quality vs GPA"
)
```

### code

```{r}
#| echo: true
#| eval: false
ggbetweenstats(
  data = df,
  x = Quality_Lecturer,
  y = GPA_num,
  type = "np",
  messages = FALSE,
  title = "Lecturer Quality vs GPA"
)
```
:::

::: callout-tip
## Observation

Students who rated lecturer quality more positively display a modest increase in the proportion of “Good” and “Excellent” GPA categories relative to those reporting lower quality perceptions. However, substantial overlap remains across groups, particularly within the dominant “Fair” category. Although an upward trend is observable, the gradient appears less pronounced than that observed for adaptation or study time. The distributional shifts suggest a positive but moderate association between perceived instructional quality and academic performance. Differences between adjacent quality levels are relatively small, indicating limited structural separation.Although statistically significant (χ² = 25.90, p \< .001), the effect size (ε² = 0.01) indicates a small practical impact. Compared with behavioral engagement and learning adaptation, lecturer quality appears to exert a weaker influence on GPA composition, though it remains directionally consistent with improved academic outcomes.
:::

## 6.3 Lecturer Support and GPA

To further examine lecturer-related influences, we analyze GPA distribution across levels of perceived lecturer support. Lecturer support is treated as an ordered Likert-scale variable, and GPA remains an ordered categorical outcome. A boxplot with non-parametric comparison is used to assess median shifts while respecting the ordinal nature of the data.

::: panel-tabset
### Plot

```{r}
#| echo: false
#| message: false
#| warning: false
ggbetweenstats(
  data = df,
  x = SupportOf_Lec,
  y = GPA_num,
  type = "np",
  messages = FALSE,
  title = "Lecturer Support vs GPA"
)
```

### Code

```{r}
#| echo: true
#| eval: false
ggbetweenstats(
  data = df,
  x = SupportOf_Lec,
  y = GPA_num,
  type = "np",
  messages = FALSE,
  title = "Lecturer Support vs GPA"
)
```
:::

::: callout-tip
## Observation

The distribution of GPA across support levels reveals a modest upward shift in central tendency as perceived support increases. Students reporting higher levels of lecturer support tend to exhibit slightly higher median GPA values and a marginally greater concentration within the upper performance categories. However, the interquartile ranges across support groups remain substantially overlapping, indicating limited structural separation.

Although the Kruskal–Wallis test suggests statistical significance, the observed effect size is small, implying that the magnitude of differentiation is modest in practical terms. Compared with learning adaptation and study time—both of which demonstrated clearer monotonic gradients—the association between lecturer support and GPA appears weaker and less structurally distinct.
:::

These findings suggest that while perceived lecturer support aligns directionally with improved academic outcomes, it likely operates in conjunction with other behavioral and institutional factors rather than serving as a primary differentiating mechanism. The overall pattern indicates a positive but limited association within the broader network of influences on academic performance.

# 7. Summary

## 7.1 Overall Key Observations

Across the examined variables, differences in academic performance are better characterized by relationship structure rather than magnitude alone. Demographic characteristics such as gender show distributional similarity, while family economic status introduces only a mild shift around central performance categories. In contrast, behavioral engagement and learning adaptation display ordered gradients, where higher levels correspond to progressively higher placement within the GPA hierarchy. Social media usage shows irregular variation without a consistent directional pattern. Lecturer-related perceptions demonstrate small but directionally consistent associations.

Overall, the results suggest that academic performance in this dataset is aligned more with engagement and adjustment processes than with static background characteristics. Rather than sharply separating groups, most factors operate by gradually redistributing students along an existing performance continuum.



## 7.2 Interpretation

These findings indicate that academic outcomes are structured through cumulative behavioral and contextual alignment rather than single dominant predictors. Variables related to active participation in learning exhibit systematic ordering effects, whereas demographic attributes mainly contribute marginal compositional differences. The analysis therefore highlights performance positioning along a continuum rather than categorical performance gaps.


## 7.3 Limitations

Several limitations should be acknowledged.

First, the dataset relies on self-reported survey responses, which may introduce reporting bias or subjective interpretation of scale categories.

Second, GPA was analyzed as an ordered categorical variable. While this preserves ordinal integrity, it reduces granularity relative to continuous GPA measures.

Third, group sizes across certain response categories are imbalanced, which may affect the stability of proportional comparisons.

Finally, the analysis is exploratory and visually driven. Although it identifies structural patterns, it does not control for confounding variables or model potential interactions among demographic, behavioral, and institutional factors.

# 8 Reference 

https://data.mendeley.com/datasets/23ppcdbmhc/1

https://www.sciencedirect.com/science/article/pii/S2352340925001702
