---
title: "Hands-on_Ex05a"
author: "Wu Yanjie"
---

# Creating Ternary Plot with R

## Overview

In this exercise, I will learn how to build a ternary plot to visualize three-part compositional data. A ternary plot is a triangle chart used to show the proportion of three components. Each side of the triangle represents one component, scaled from 0 to 1.

The following packages will be used:

tidyverse (dplyr): Use mutate() to create three new measures

ggtern: Build a static ternary plot

plotly: Create an interactive ternary plot

## Installing and launching R packages

For this exercise, two main R packages will be used in this hands-on exercise, they are:

ggtern, a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots. Plotly R, an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library, plotly.js . The plotly R libary contains the ggplotly function, which will convert ggplot2 figures into a Plotly object.

```{r}
pacman::p_load(plotly, ggtern, tidyverse)
```

## Importing Data

```{r}
#Reading the data into R environment
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

## Preparing the Data

Next, use the mutate() function of dplyr package to derive three new measures, namely: young, active, and old.

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)
```

## Plotting Ternary Diagram with R

Plotting a static ternary diagram Use ggtern() function of ggtern package to create a simple ternary plot.

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

Plotting an interative ternary diagram The code below create an interactive ternary plot using plot_ly() function of Plotly R.

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```
